Visualization-Focused Continuous Monitoring of Evolving Flask \& Python-Based Web Services

### 2018-04-09 paper.tex
one more tiny detail
submitted version of the paper

% An Agile Analytics Platform for Visualizing Evolving Flask-Based Python Web Services
% A Dashboard for the Agile Monitoring of Evolving Web Services: For the Little Man
% A Dashboard for Monitoring Flask \& Python-Based Web Service Evolution
% Monitoring the Evolution of Flask \& Python-Based Web Services
% A Dashboard for Monitoring the Evolution of Flask \& Python-Based Web Services
\title{Monitoring the Evolution of Flask \& Python-Based Web Services}

@@

In the previous sections we discussed our proposal for a monitoring dashboard for Flask-based web applications as the means for providing a low effort and flexible analytics solution to projects on a budget. The emphasis is in allowing application developers to gain insight into how the performance and utilization of their services evolves together with the application itself. For this purpose the \tool allows for integration with both \git and the Travis CI platform. We also discussed more advanced features of the dashboard in the form of grouping results by user groups, and provided evidence towards the use of unit testing-sourced synthetic loads as a potential limited form of performance prediction. Finally we measured the overhead of the proposed solution within acceptable by the case study application owner limits. Section~\ref{sec:discussion} which discusses what we perceive as limitations of our proposal and also potential challenges for other similar tools, effectively acts also as a roadmap for further development, with the ultimate goal of increasing the adoption of the \tool by as many projects as possible. This will allow us to evaluate our proposal in depth in the future.

@@

\bibliographystyle{abbrv}
\bibliography{paper}

>>>>

% An Agile Analytics Platform for Visualizing Evolving Flask-Based Python Web Services
% A Dashboard for the Agile Monitoring of Evolving Web Services: For the Little Man
% A Dashboard for Monitoring Flask \& Python-Based Web Service Evolution
% Monitoring the Evolution of Flask \& Python-Based Web Services
% A Dashboard for Monitoring the Evolution of Flask \& Python-Based Web Services
% Continuously Monitoring the Evolution of Flask \& Python-Based Web Services
% A Dashboard for Continuously Monitoring the Evolution of Flask \& Python-Based Web Services
% something about visualizing
% Continuously Monitoring Through Visualization the Evolution of Flask \& Python-Based Web Services
% Visual Perspectives for Continuously Monitoring the Evolution of Flask \& Python-Based Web Services
\title{Visualization-Focused Continuous Monitoring of Evolving Flask \& Python-Based Web Services}

@@

In this paper we discussed our proposal for a monitoring dashboard for Flask-based web applications as the means for providing a low effort analytics solution. The emphasis is in allowing application developers to gain insight into how the performance and utilization of their services evolves together with the application itself. \tool allows for integration with both \git and the Travis CI platform. We also discussed more advanced features of the dashboard and provided evidence towards the use of unit testing-sourced synthetic loads as a potential limited form of performance prediction. Finally we measured the overhead of the proposed solution within acceptable by the case study application owner limits. Section~\ref{sec:discussion} which discusses what we perceive as limitations of our proposal and also potential challenges for other similar tools, effectively acts also as a roadmap for further development, with the ultimate goal of increasing the adoption of the \tool by as many projects as possible. This will allow us to evaluate our proposal in depth in the future.

@@

\newpage
\bibliographystyle{abbrv}
\bibliography{paper}

### 2018-04-09 related-work.tex
submitted version of the paper

A rich body of work exists that studies the evolution and maintenance of APIs from the perspective of source code evolution \cite{dig2006apis, hora2015developers, hora2015apiwave}. However, since services are increasingly being used across a variety of systems in which adaptability, flexibility, and environmental awareness are essential~\cite{pernici2016monitoring} in our work we focus on the analysis of the runtime aspects of service APIs. Our work is thus, more related to the work on dynamic analysis of software systems; however, most of this work has been done in order to support the reverse engineering of software systems\cite{Corn09-dynamic}.

In this context, our work falls within the server-side run-time monitoring of services ~\cite{ghezzi2007run}. While we don't implement the more advanced features of related monitoring solutions like QoS policies driving the monitoring, it presents nevertheless an easy to use approach towards improving the performance of web applications.

There is a long tradition of using visualization for gaining insight into software performance. Tools like Jinsight \cite{Pauw02a} and Web Services Navigator \cite{Pauw05} pioneered such an approach for Java and for Web Services that communicate with SOAP messages. Both have an ``omniscient'' view of the services / objects and their interactions. As opposed to them, in our work we present an analytics platform which focuses on monitoring a single Python web service from its own point of view.

Kieker -- has been presented by Rohr et al \cite{Rohr08} as a system targeted at monitoring the performance of Java services. The system provides a better performance in terms of overhead, but does not support integration with git and multi-version analysis as the system presented here.

>>>>

A rich body of work exists that studies the evolution and maintenance of APIs from the perspective of source code evolution \cite{dig2006apis, hora2015developers, hora2015apiwave}. However, since services are increasingly being used across a variety of systems in which adaptability, flexibility, and environmental awareness are essential~\cite{pernici2016monitoring} in our work we focus on the analysis of the runtime aspects of service APIs. Our work is thus, more related to the work on dynamic analysis of software systems; however, most of this work has been done in order to support the reverse engineering and enhance the understanding of software systems\cite{Corn09-dynamic}.

In this context, our work falls within the server-side run-time monitoring of services ~\cite{ghezzi2007run}. 
% While we don't implement the more advanced features of related monitoring solutions like QoS policies driving the monitoring, it presents nevertheless an easy to use approach towards improving the performance of web applications.
There is a long tradition of using visualization for gaining insight into software performance. Tools like Jinsight \cite{Pauw02a} and Web Services Navigator \cite{Pauw05} pioneered such an approach for Java and for Web Services that communicate with SOAP messages. More recent tools present live visualizations for monitoring multiple interacting software systems utilizing the city metaphor for each software system \cite{fittkau2013live}. All these approaches have an ``omniscient'' view of the services / objects and their interactions. As opposed to them, in our work we present an analytics platform which focuses on monitoring a single Python web service from its own point of view.

Kieker -- has been presented by Rohr et al \cite{Rohr08} as a system targeted at monitoring the performance of Java services. The system provides better performance in terms of overhead with respect to ours, but does not support integration with git and multi-version analysis as the system presented here.

Similar to our discussion on pre-emptive monitoring, Cito et al. \cite{cito2015runtime} argue for bringing performance measurements back in the IDE to provide the developers with such information during the development process.
